Safe and Robust Formal Synthesis of Digital Controllers for Continuous Plants

### 2018-03-08 automatica/paper.tex

The linear feedback controller is characterized by a gain matrix
$\mat{K}$ that influences the closed-loop dynamics
(Fig.~\ref{fig:digitalsystem}) such that
\begin{equation*}
\vec{u}_k=\mat{K}\vec{x}_k \Rightarrow \vec{x}_{k+1}=(\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_k
\end{equation*}
Which corresponds to a closed loop system without inputs. Later, we will address the issue of
variable quantization, as introduced by the ADC/DAC conversion blocks.

@@

We recall a technique called \emph{Abstract Acceleration}~\cite{JSS14,cattaruzza2015unbounded}, 
which will be key for abstraction-based synthesis.  
%\note{This clashes with the model introduced above, with inputs - motivate why? I've added the motivation above by introducing the closed  loop}
Given a model of an iterative program with dynamics
%
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k \text{ where } \vec{x}_k \in \mathbb{R}^p, \mat{A} \in \mathbb{R}^{p \times p} \text{ and } k \in \mathbb N_0,
\end{equation}
%
we want to find all possible states visited over an infinite time horizon
starting at an initial set $X_0$.  
This set is known as the \emph{reach tube} of
the model:
%
\begin{equation}
\hat{X} = \{ \vec{x}_k \mid k \in \mathbb N_0, \vec{x}_0 \in X_0 \text{ and } \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
%
Using Abstract Acceleration we
transform this problem into a one-step solution.  We first change the
problem statement by means of acceleration into
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 \mid k \in N_0 \text{ and } \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and further define new model semantics and corresponding set: 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model \cite{JSS14}.  All we need
to do is define the nature of $\mathcal{A}X_0$.

Let $\mat{S}\mat{J}\mat{S}^{-1}=\mat{A}$ be the eigen-decomposition of
$\mat{A}$, where $\mat{S}$ is the set of generalised eigenvectors of
$\mat{A}$ and $\mat{J}$ is the Jordan Form containing the eigenvalues of
$\mat{A}$ in its main diagonal.  We correspondingly define
$\mathcal{J}=\mat{S}^{-1}\mathcal{A}\mat{S}$, which is such that
$\mathcal{J} \supseteq \bigcup_{k \in [0\ \infty)} \mat{J}^k$.

Since abstract acceleration operates in the eigenspace, wemust deal with some preliminary transformations.
Given a set $X=\{ \vec{x} \mid \mat{C}\vec{x} \leq \vec{d}\}$,
%\note{Why are we interested in this set? What does this set represent?}
the linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to the set 
%
\begin{equation}
X'= \mat{S}^{-1}X = \{\vec{x}' \mid \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
\label{eq:eigenInit}
\end{equation}
%
Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
transform equation \eqref{eq:aa_reachtube} into: 
%
\begin{equation}
\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', 
\end{equation}
%
where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
tableau defined by $\mathcal{J}$ and objective functions derived from each
vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
\ref{alg:findReachTube} below).  We note, however, that if we wish to use
numerical analysis in the decomposition of $\mat{A}$, the Tableau in
$\mathcal{J}$ contains small intervals given by the errors, which cannot be
processed efficiently by a regular simplex and require a new procedure.

The calculation of the abstract reach tube via abstract acceleration is
encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
follows (its steps are denoted in parentheses):
%
\begin{enumerate}
%
\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%
\item Then we restore soundness to the results using the methods described in \cite{cattaruzza2017sound} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}).
%
\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%
\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}). 
%
\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%
\item Extract the vertices of the eigen-polyhedron $X_0'$ using the algorithms described in \cite{cattaruzza2015unbounded,cattaruzza2017sound} (line \ref{alg:getVertices}). 
%
\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
that over-approximates the image of the reach tube in the eigenspace (line
\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
functions using the abstract dynamics as a simplex Tableau, and via a sound
simplex described in \cite{cattaruzza2017sound}.  The objective functions are
defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
template directions, such that
%
\begin{equation*}
\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j \text{ with } \vec{v}_i \in V_0 \text{ and } \vec{t}_j \in T,
\end{equation*}
%
where $T$ is the set of template directions, and $\circ$ denotes a
component-wise multiplication yielding a vector.
%
\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%
\end{enumerate}

\begin{algorithm}[]
\small
\begin{algorithmic}[1]
\Function{$\mathit{findAbstractReachTube}()$}{}
\State $\hat{\mat{J}}=$ calculateEigenvalues($ \mat{A}$)
\label{alg:calcEigenvalues}
\State $\hat{\mat{S}}=$ calculateEigenvectors($ \mat{A}$)
\label{alg:calcEigenvectors}
\State $[\mat{J}]=$ soundifyEigenvalues($\mat{A},\mat{J},\mat{S}$)
\label{alg:soundEigenvalues}
\State $[\mat{S}]=$ soundifyEigenvectors($\mat{A},[\mat{J}],\mat{S}$)
\label{alg:soundEigenvectors}
\State $[\mat{S}]^{-1}=$ calculateInverse($[\mat{S}]$)
\label{alg:inverse}
\State $\mathcal{J}=getAbstractDynamics([\mat{J}])$
\label{alg:aa}
\State $[X_0']=$ transformInitialSpace($X_0,\mat{S}$)
\label{alg:eigeninit}
\State $[V_0]=$ getVertices($[X_0']$)
\label{alg:getVertices}
\State ${\hat{X}'^\sharp}=$ getReachTube($\mathcal{J},[V_0]$)
\label{alg:getReachTube}
\State ${\hat{X}^\sharp}=$ transformReachTube(${\hat{X}'^\sharp},[\mat{S}]^{-1}$)
\label{alg:getNormReachTube}
\EndFunction
\end{algorithmic}
\caption{Calculation of Abstract Reach Tube Using Abstract Acceleration}
\label{alg:findReachTube}
\end{algorithm}

In the case of Discrete dynamics, support functions can be found by
evaluating hyperplanes of consecutive exponents of the eigenvalues using
differences.  In the case of continuous time, this translates to calculating
derivatives.  Once again, when the sets are known to be convex, we may
calculate the normal vector to this derivative and evaluate its support
function on the selected iteration.  The direction of the vector is
determined by evaluating the support function of another point in the curve,
which is known to be smaller than that of the selected point.  The
corresponding derivatives are:
%
\begin{enumerate}
%
\item for real eigenvalues: we first calculate the slope of the tangent
between the two progressions: $\frac{d f(k)}{d g(k)}=\frac{f'(k)}{g'(k)}$. 
Hence $\frac{d \lambda_1^k}{d \lambda_2^k}=\frac{log(\lambda_1)
\lambda_1^k}{log(\lambda_2) \lambda_2^k}$.  From this we know that the polar
angle of the support vector is $\psi = \tan^{-1}(\frac{log(\lambda_1)
\lambda_1^k}{log(\lambda_2) \lambda_2^k})+\frac{\pi}{2}$
%
\item for conjugate eigenvalue pairs: the pair forms a logarithmic spiral. 
The tangent of a spiral with polar equation $r=a e^{b\theta}$ will have a
derivative $\frac{d r}{d \theta}=b a e^{b\theta}=b r$, hence the angle of
the vector normal to the curve is $\tan^{-1}(\frac{r}{\frac{d r}{d
\theta}})+\frac{\pi}{2}=\tan^{-1}(\frac{1}{b})+\frac{\pi}{2}$
%
\item for Jordan blocks with $gm>1$: in the case of real eigenvalues, we
follow the same procedure as in the real case, but considering that the
equation for each eigenvalue is $\frac{k^j}{j!}e^{\lambda k}$ where $j$ is
the index of the term within the Jordan block (note that when $j=0$ the term
is the same as for real eigenvalues above).  The derivatives f'(k),g'(k) are
in the form $(j+k*log(\lambda))\frac{k^{j-1}}{j!}e^{\lambda k}$.
%
\item for any combination of complex eigenvalues we use the above equations
using the envelope of their magnitude (ie $|\lambda|$).
%
\end{enumerate}

@@

>>>>

Later, we will address the issue of variable quantization, as introduced by the ADC/DAC conversion blocks.
Models \eqref{eq:ode1} and \eqref{eq:plant} depend on external non-determinism in the form of input signals $u (t)$ and  $u_k$, respectively. 
Feedback architectures can be employed to manipulate the properties and behaviors of the continuous process (the plant).   
We are interested in the synthesis of digital feedback control algorithms, 
as implemented on Field-Programmable Gate Arrays or Digital Signal Processors. 
The most basic feedback architecture is the state feedback one, 
where the control action $u_k$ (notice we work with the discretized signal) is computed by: 
%
\begin{equation*}
%\label{eq:controlaction}
u_k = r_{k} - K x_k. 
\end{equation*}

where $K \in \mathbb{R}^{m \times n}$ is a state-feedback gain matrix, 
and $r_{k}$ is a reference signal (again digital). We will assume $r_k=0$, thus 
\begin{equation}
\vec{x}_{k+1}=(\mat{A}_d-\mat{B}_d\mat{K})\vec{x}_k
\end{equation}
Which corresponds to a closed loop system without inputs.

@@

@@

We recall a technique called \emph{Abstract Acceleration}~\cite{JSS14,cattaruzza2015unbounded}, 
which will be key for abstraction-based synthesis.  
%\note{This clashes with the model introduced above, with inputs - motivate why? I've added the motivation above by introducing the closed  loop}
Given a model of an iterative program with dynamics
%
\begin{equation}
\vec{x}_{k+1}=\mat{A}\vec{x}_k \text{ where } \vec{x}_k \in \mathbb{R}^p, \mat{A} \in \mathbb{R}^{p \times p} \text{ and } k \in \mathbb N_0,
\end{equation}
%
we want to find all possible states visited over an infinite time horizon
starting at an initial set $X_0$.  
This set is known as the \emph{reach tube} of
the model:
%
\begin{equation}
\hat{X} = \{ \vec{x}_k \mid k \in \mathbb N_0, \vec{x}_0 \in X_0 \text{ and } \vec{x}_{k+1}=\mat{A}\vec{x}_k\}.
\end{equation}
%
Using Abstract Acceleration we
transform this problem into a one-step solution.  We first change the
problem statement by means of acceleration into
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A}^k\vec{x}_0 \mid k \in N_0 \text{ and } \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and further define new model semantics and corresponding set: 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model \cite{JSS14}.  
In this work we use an extension of the abstract acceleration model for
continuous time. For this reason we must first look at the accelerationm of continuous time dynamics.

%All we need to do is define the nature of $\mathcal{A}X_0$.
%Let $\mat{S}\mat{J}\mat{S}^{-1}=\mat{A}$ be the eigen-decomposition of
%$\mat{A}$, where $\mat{S}$ is the set of generalised eigenvectors of
%$\mat{A}$ and $\mat{J}$ is the Jordan Form containing the eigenvalues of
%$\mat{A}$ in its main diagonal.  We correspondingly define
%$\mathcal{J}=\mat{S}^{-1}\mathcal{A}\mat{S}$, which is such that
%$\mathcal{J} \supseteq \bigcup_{k \in [0\ \infty)} \mat{J}^k$.
%Since abstract acceleration operates in the eigenspace, wemust deal with some preliminary transformations.
%Given a set $X=\{ \vec{x} \mid \mat{C}\vec{x} \leq \vec{d}\}$,
%the linear transformation $\vec{x}'=\mat{S}^{-1}\vec{x}$ leads to the set 
%X'= \mat{S}^{-1}X = \{\vec{x}' \mid \mat{C}\mat{S}\vec{x}'\leq \vec{d}\}.  
%\label{eq:eigenInit}
%Analogously, let $X_0'=\mat{S}^{-1}X_0$ and
%${\hat{X}'^\sharp}=\mat{S}^{-1}{\hat{X}^\sharp}$ be the initial set and the
%abstract reach tube mapped onto the eigenspace of matrix $\mat{A}$.  We may
%transform equation \eqref{eq:aa_reachtube} into: 
%\hat{X}' \subseteq {\hat{X}'^\sharp} = \mathcal{J}X_0', 
%where $\hat{X}'=\mat{S}^{-1}\hat{X}$ and $\mathcal{J}$ is a convex
%polyhedron representing restrictions on the eigenvalues of $\mat{A}$. 
%$\hat{X}'^\sharp$ is calculated by applying a simplex algorithm with a
%tableau defined by $\mathcal{J}$ and objective functions derived from each
%vertex of $X_0'$ (step \eqref{alg:getReachTube} in Algorithm
%\ref{alg:findReachTube} below).  We note, however, that if we wish to use
%numerical analysis in the decomposition of $\mat{A}$, the Tableau in
%$\mathcal{J}$ contains small intervals given by the errors, which cannot be
%processed efficiently by a regular simplex and require a new procedure.
%The calculation of the abstract reach tube via abstract acceleration is
%encompassed in Algorithm~\ref{alg:findReachTube}, which can be summarised as
%follows (its steps are denoted in parentheses):
%\item We perform unsound eigen-decomposition using an existing algebra package (lines \ref{alg:calcEigenvalues}-\ref{alg:calcEigenvectors}).
%\item Then we restore soundness to the results using the methods described in \cite{cattaruzza2017sound} (lines \ref{alg:soundEigenvalues}-\ref{alg:soundEigenvectors}).
%\item The inverse of the matrix of eigenvectors is calculated after soundness is restored, using interval arithmetic in order to ensure its soundness (line \ref{alg:inverse}).
%\item The abstract dynamics are obtained by evaluating the convex hull of all powers of eigenvalues up to the desired number of iterations as described in \cite{cattaruzza2015unbounded} (line \ref{alg:aa}). 
%\item Using Equation \eqref{eq:eigenInit}, transform the initial state into the eigenspace (line \ref{alg:eigeninit}).
%\item Extract the vertices of the eigen-polyhedron $X_0'$ using the algorithms described in \cite{cattaruzza2015unbounded,cattaruzza2017sound} (line \ref{alg:getVertices}). 
%\item Calculate the mapped \emph{abstract reach tube} ${\hat{X}'^\sharp}$
%that over-approximates the image of the reach tube in the eigenspace (line
%\ref{alg:getReachTube}).  This is achieved by evaluating a set of objective
%functions using the abstract dynamics as a simplex Tableau, and via a sound
%simplex described in \cite{cattaruzza2017sound}.  The objective functions are
%defined by the vertices of $X'_0$, denote as $V_0$, and by the desired
%template directions, such that
%\begin{equation*}
%\vec{w}_{ij}= \vec{v}_i \circ \vec{t}_j \text{ with } \vec{v}_i \in V_0 \text{ and } \vec{t}_j \in T,
%\end{equation*}
%where $T$ is the set of template directions, and $\circ$ denotes a
%component-wise multiplication yielding a vector.
%\item Using Equation~\eqref{eq:eigenInit}, find the reach tube
%$\hat{X}^\sharp=\mat{S}{\hat{X}'^\sharp}$ (line \ref{alg:getNormReachTube}).
%\begin{algorithm}[]
%\small
%\begin{algorithmic}[1]
%\Function{$\mathit{findAbstractReachTube}()$}{}
%\State $\hat{\mat{J}}=$ calculateEigenvalues($ \mat{A}$)
%\label{alg:calcEigenvalues}
%\State $\hat{\mat{S}}=$ calculateEigenvectors($ \mat{A}$)
%\label{alg:calcEigenvectors}
%\State $[\mat{J}]=$ soundifyEigenvalues($\mat{A},\mat{J},\mat{S}$)
%\label{alg:soundEigenvalues}
%\State $[\mat{S}]=$ soundifyEigenvectors($\mat{A},[\mat{J}],\mat{S}$)
%\label{alg:soundEigenvectors}
%\State $[\mat{S}]^{-1}=$ calculateInverse($[\mat{S}]$)
%\label{alg:inverse}
%\State $\mathcal{J}=getAbstractDynamics([\mat{J}])$
%\label{alg:aa}
%\State $[X_0']=$ transformInitialSpace($X_0,\mat{S}$)
%\label{alg:eigeninit}
%\State $[V_0]=$ getVertices($[X_0']$)
%\label{alg:getVertices}
%\State ${\hat{X}'^\sharp}=$ getReachTube($\mathcal{J},[V_0]$)
%\label{alg:getReachTube}
%\State ${\hat{X}^\sharp}=$ transformReachTube(${\hat{X}'^\sharp},[\mat{S}]^{-1}$)
%\label{alg:getNormReachTube}
%\EndFunction
%\end{algorithmic}
%\caption{Calculation of Abstract Reach Tube Using Abstract Acceleration}
%\label{alg:findReachTube}
%\end{algorithm}

Now that we have defined the dynamics of continuous time, we can use abstract acceleration techniques to evaluate
the above model such that
\begin{equation}\label{eq:aa_cont_amatrix}
\bigcup_{T \in [0, T_s], k \in N_0} (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k \subseteq \mathcal{A}.
\end{equation}

Abstract acceleration works by finding support functions of the set that encompasses the matrices $\mat{A}^k$.
Support functions are scalars that measure the distance of a hyperplane tangent to the set boundary to the origin.
In the case of continuous spaces, this translates to calculating the derivatives of the envelope in order to find the direction of the support vector (normal to the hyperplane), and calculating the distance to the corresponding point where the hyperplane touches the set (at which the derivative was calculated).
The direction of the vector is determined by evaluating the support function of another point in the curve,
which is known to be smaller than that of the selected point, and the chosen hyperplanes are aligned with the eigenvectors, hence we only look at eigenvalue pairs to calculate the derivatives~\cite{cattaruzza2015unbounded}.
The corresponding derivatives are:
%
\begin{enumerate}
%
\item for real eigenvalues: we first calculate the slope of the tangent
between the two progressions: $\frac{d f(k)}{d g(k)}=\frac{f'(k)}{g'(k)}$. 
Hence $\frac{d \lambda_1^k}{d \lambda_2^k}=\frac{log(\lambda_1)
\lambda_1^k}{log(\lambda_2) \lambda_2^k}$.  From this we know that the polar
angle of the support vector is $\psi = \tan^{-1}(\frac{log(\lambda_1)
\lambda_1^k}{log(\lambda_2) \lambda_2^k})+\frac{\pi}{2}$
%
\item for conjugate eigenvalue pairs: the pair forms a logarithmic spiral. 
The tangent of a spiral with polar equation $r=a e^{b\theta}$ will have a
derivative $\frac{d r}{d \theta}=b a e^{b\theta}=b r$, hence the angle of
the vector normal to the curve is $\tan^{-1}(\frac{r}{\frac{d r}{d
\theta}})+\frac{\pi}{2}=\tan^{-1}(\frac{1}{b})+\frac{\pi}{2}$
%
\item for Jordan blocks with $gm>1$: in the case of real eigenvalues, we
follow the same procedure as in the real case, but considering that the
equation for each eigenvalue is $\frac{k^j}{j!}e^{\lambda k}$ where $j$ is
the index of the term within the Jordan block (note that when $j=0$ the term
is the same as for real eigenvalues above).  The derivatives f'(k),g'(k) are
in the form $(j+k*log(\lambda))\frac{k^{j-1}}{j!}e^{\lambda k}$.
%
\item for any combination of complex eigenvalues we use the above equations
using the envelope of their magnitude (ie $|\lambda|$).
%
\end{enumerate}
